SMT-Boosted Security Types and Correctness for Low-Level MPC

### 2025-11-25 toplas26/model.tex
checkpoint

Pmfs map \emph{realizations} of variables in a joint
distribution to values in $[0..1]$. For convenience and following
\cite{barthe2019probabilistic,skalka-near-ppdp24} we use memories to
represent realizations, so for example given a pmf $\pmf$ over
variables $\{ \sx{x}{1}, \mx{y}{2} \}$ we write $\pmf(\{
\elab{\secret{x}}{1} \mapsto 0, \elab{\mesg{y}}{2} \mapsto 1 \})$ to
denote the (joint) probability that $\elab{\secret{x}}{1} = 0 \wedge
\elab{\mesg{y}}{2} = 1$. Recall from Section
\ref{section-lang-semantics} that $\uplus$ denotes the combination of
memories, so for example $\{ \elab{\secret{x}}{1} \mapsto 0\} \uplus
\{\elab{\mesg{y}}{2} \mapsto 1 \} = \{ \elab{\secret{x}}{1} \mapsto 0,
\elab{\mesg{y}}{2} \mapsto 1 \}$.
\begin{definition}
  A \emph{probability mass function} $\pmf$ is a function
  mapping memories in $\mems(X)$ (the \emph{domain} of $\pmf$), to
  values in $\mathbb{R}$ such that:
  $$
  \sum_{\store \in \mems(X)} \pmf(\store) \  = \ 1
  $$
  %The \emph{support} of a distribution is the set of realizations
  %of random variables with non-zero probability:
  %$\support(\pmf) \defeq \{ (v_1,\ldots,v_n) \mid
  %\pmf(\{ x_1 \mapsto v_1, \ldots, x_n \mapsto v_n\}) > 0 \} $
  We write $\rvs(\pmf) = X$ to denote the variables
  over which $\pmf$ is defined. 
\end{definition}
%To recover succinct and familiar notation, we may omit the domain of a
%distribution when it is clear from an application context-
%i.e., we allow the following sugaring:
%$$
%\pdf{}(\store) \defeq \pdf{\dom(\store)}(\store)
%$$
Now, we can define a notion of marginal and conditional
distributions as follows, which are standard for discrete
probability mass functions. 
\begin{definition}
  Given $\pmf$ with $\rvs(\pmf) = X_2$, the \emph{marginal distribution}
  of variables $X_1 \subseteq X_2$ in $\pmf$ is denoted $\margd{\pmf}{X_1}$ and defined as follows:
  $$
  \forall \store \in \mems(X_1) \quad . \quad \margd{\pmf}{X_1}(\store) \defeq
  \sum_{\store' \in \mems(X_2-X_1)} \pmf(\store \uplus \store')
  $$
\end{definition}

>>>>

Pmfs map \emph{realizations} of variables in a joint
distribution to values in $[0..1]$. For convenience and following
\cite{barthe2019probabilistic,skalka-near-ppdp24} we use memories to
represent realizations, so for example given a pmf $\pmf$ over
variables $\{ \sx{x}{1}, \mx{y}{2} \}$ we write $\pmf(\{
\elab{\secret{x}}{1} \mapsto 0, \elab{\mesg{y}}{2} \mapsto 1 \})$ to
denote the (joint) probability that $\elab{\secret{x}}{1} = 0 \wedge
\elab{\mesg{y}}{2} = 1$. Recall from Section
\ref{section-lang-semantics} that $\uplus$ denotes the combination of
memories, so for example $\{ \elab{\secret{x}}{1} \mapsto 0\} \uplus
\{\elab{\mesg{y}}{2} \mapsto 1 \} = \{ \elab{\secret{x}}{1} \mapsto 0,
\elab{\mesg{y}}{2} \mapsto 1 \}$.
\begin{definition}
  A \emph{probability mass function} $\pmf$ over variables $X$
  (where we define $\rvs(\pmf) \defeq X$) is a function
  with domain $\mems(X)$ and range $\mathbb{R}$ such that:
  $$
  \sum_{\store \in \mems(X)} \pmf(\store) \  = \ 1
  $$
  %The \emph{support} of a distribution is the set of realizations
  %of random variables with non-zero probability:
  %$\support(\pmf) \defeq \{ (v_1,\ldots,v_n) \mid
  %\pmf(\{ x_1 \mapsto v_1, \ldots, x_n \mapsto v_n\}) > 0 \} $
  %We write $\rvs(\pmf) = X$ to denote the variables
  %over which $\pmf$ is defined. 
\end{definition}
%To recover succinct and familiar notation, we may omit the domain of a
%distribution when it is clear from an application context-
%i.e., we allow the following sugaring:
%$$
%\pdf{}(\store) \defeq \pdf{\dom(\store)}(\store)
%$$
Now, we can define a notion of marginal and conditional
distributions as follows, which are standard for discrete
probability mass functions. 
\begin{definition}
  Given $\pmf$ with $\rvs(\pmf) = X_2$, the \emph{marginal distribution}
  of variables $X_1 \subseteq X_2$ in $\pmf$ is denoted $\margd{\pmf}{X_1}$ and defined as follows:
  $$
  \forall \store \in \mems(X_1) \quad . \quad \margd{\pmf}{X_1}(\store) \defeq
  \sum_{\store' \in \mems(X_2-X_1)} \pmf(\store \uplus \store')
  $$
\end{definition}

### 2025-11-25 toplas26/sty.tex
checkpoint

Share type soundness (Theorem \ref{theorem-sty}) means that share
typing enforces noninterference modulo output. To prove this we need
to assert probabilistic properties in share typings, including for
functions of protocol variables such as reconstruction of shares. So
we make standard extensions of notation for probability distributions
to express probabilities, and conditional probabilities, of functions
of program variables $\phi$. For this extension we posit that
memories can incorporate variables $x_\phi$ that assign valuations
to functions $\phi$.
\begin{definition}
  We let $\Phi$ range over sets of SMT formulas $\phi$, and define
  $
  \mems(\Phi) \defeq \{ \store \mid \dom(\store) = \setit{\ x_{\phi}\ \mid\ \phi \in \Phi\ } \}
  $.
  Given pmf $\pmf$ and $\Phi$ with $\dom(\pmf) = \bigcup_{\phi \in \Phi} \dom(\phi)$,
  for all $\store \in \mems(\Phi)$ define:
%$$
%\pmf(\phi = v) \defeq \sum_{\store \in \mems(X) \ \text{with}\ \store(\phi) = v} \pmf(\store)
%$$
  $$
  \pmf(\store) \defeq \sum_{\store' \in \mems(X)}
  \begin{cases}
    \pmf(\store') \ \text{if}\ \forall \phi \in \Phi . \store'(\phi) = \store(x_\phi)  \\
    0 \ \text{otherwise}
  \end{cases}
  $$
\end{definition}

\begin{definition}
  Given $\pmf$, the \emph{conditional distribution}
  of $X$ given $\Phi$ where $\bigcup_{\phi \in \Phi} \dom(\phi) \cup X = \dom(\pmf)$ 
  is denoted $\condd{\pmf}{X}{\Phi}$ and defined as follows: 
  $$
  \forall \store_1 \in \mems(X)\ .\ \forall \store_2 \in \mems(\Phi) \ .\ 
  \condd{\pmf}{X}{\Phi}(\store_1 \uplus \store_2) \defeq
  \begin{cases}
    0 \text{\ if\ } \pmf(\store_2) = 0\\
    \pmf(\store_1 \uplus \store_2) / \pmf(\store_2) \text{\ otherwise}
  \end{cases}
  $$
\end{definition}

>>>>

Share type soundness (Theorem \ref{theorem-sty}) means that share
typing enforces noninterference modulo output. To prove this we need
to assert probabilistic properties in share typings, including for
functions of protocol variables such as reconstruction of shares. So
we make standard extensions of notation for probability distributions
to express probabilities, and conditional probabilities, of functions
of program variables $\phi$. For this extension we posit that
memories can incorporate variables $x_\phi$ that assign valuations
to functions $\phi$.
\begin{definition}
  We let $\Phi$ range over sets of SMT formulas $\phi$, and define
  $
  \mems(\Phi) \defeq \{ \store \mid \dom(\store) = \setit{\ x_{\phi}\ \mid\ \phi \in \Phi\ } \}
  $.
  Given pmf $\pmf$ and $\Phi$ with $\rvs(\pmf) = \bigcup_{\phi \in \Phi} \dom(\phi)$,
  for all $\store \in \mems(\Phi)$ define:
%$$
%\pmf(\phi = v) \defeq \sum_{\store \in \mems(X) \ \text{with}\ \store(\phi) = v} \pmf(\store)
%$$
  $$
  \pmf(\store) \defeq \sum_{\store' \in \mems(X)}
  \begin{cases}
    \pmf(\store') \ \text{if}\ \forall \phi \in \Phi . \store'(\phi) = \store(x_\phi)  \\
    0 \ \text{otherwise}
  \end{cases}
  $$
\end{definition}

\begin{definition}
  Given $\pmf$, the \emph{conditional distribution}
  of $X$ given $\Phi$ where $\bigcup_{\phi \in \Phi} \dom(\phi) \cup X = \rvs(\pmf)$ 
  is denoted $\condd{\pmf}{X}{\Phi}$ and defined as follows: 
  $$
  \forall \store_1 \in \mems(X)\ .\ \forall \store_2 \in \mems(\Phi) \ .\ 
  \condd{\pmf}{X}{\Phi}(\store_1 \uplus \store_2) \defeq
  \begin{cases}
    0 \text{\ if\ } \pmf(\store_2) = 0\\
    \pmf(\store_1 \uplus \store_2) / \pmf(\store_2) \text{\ otherwise}
  \end{cases}
  $$
\end{definition}
